\vspace{-.2in} Squint: Lossy Hierarchical Compression on Symbol Streams \vspace{-.1in}

### 2012-12-12 tex/paper.tex
Number numbered things

\begin{itemize}
  \item Infer a CFG representation of a stream (we use \Sequitur).
  \item Simplify the CFG, introducing loss.
  \item Encode and output the CFG.
\end{itemize}

@@

Overview

@@

@@

For our first suite of tests, we compare the effectiveness of the
\Similarity and \Cluster algorithms. We might expect \Similarity to
achieve better quality at the same compression ratio since it updates the state
of the grammar between each variable replacement, on the other hand since it
must update the grammar it has the potential to run much more slowly than
\Cluster. The running times for just \Sequitur aren't reported since its linear running
time is negligible compared to our algorithms.

In table \ref{simclus_r_t} we report the results on running \Sequitur and
\Cluster, without the grammar reductions, on the rep-n datasets. In table
\ref{simclus_g_t} we do the same for the gen-n datasets.
In these tests
we wanted to evaluate the relative fidelity, compression ratios, and runtime 
of the two algorithms.

In the table, Time is the
wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
Ratio is the relative size of the grammar produced compared with \Sequitur,
where the size of the grammar is the sum of the lengths of the rhs's of the
rules. Levdistance is the levenshtein distance between the output of the
lossified grammar and the original input string, it is a kind of measure of the
fidelity of our algorithms but doesn't take into account how well patterns were
preserved. For the rep-n tests, proportion is the proportion of the final
string produced that was `*'. The closer to 1/30~0.033, the more accurate the
lossifier was in recreating the original texture of the input.
The columns of the data are prefixed by S or T depending on whether they are
measurements for \Similarity or \Cluster respectively.

In both datasets, the 
compression ratios achieved for both algorithms are roughly equal given the
same value for $\epsilon$, so it is fair to compare the algorithms for the same
$\epsilon$. 
From plotting the data, we see that runtime of the algorithms are roughly
polynomial in the size of the data, so we can perform a linear regression on
the log-log transform of the data to find the exponents $\alpha$ for the
runtime $r(n)=n^\alpha$ where $n$ is the length of the input string. In the
chart below, all of the linear regressions had $R^2$ values of over $0.98$.

\begin{tabular}{l|l|l|l|l}
Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
Exponent & $1.859$ & $2.026$ & $1.306$ & $1.305$
\end{tabular}

For both values of $\epsilon$, The runtime for the
clustering algorithm is nearly linear, while
the similarity algorithm is roughly quadratic. On this dataset, the leveshtein
distances were similar for both algorithms but fairly high since the length of
the final string could change drastically.  The proportion of '*'s in the input
was usually the right order of magnitude, with both algorithms making big
mistakes occassionally. In most cases, the two algorithms maintained similarly
acceptable fidelity levels in recreating the texture of the original rep-n
input file.

As mentioned earlier,
the actual compression ratio was respectable, with both \Similarity and \Cluster
coming in at nearly half the size of the \Sequitur CFG. 
However, setting higher values
for $\epsilon$ had diminishing returns for compression for larger
file sizes, at the cost of drastically lowering the fidelity of the output.

Just as with the rep-n data, log-log regression tests yield the following
exponents for the runtime on the gen-n data, again with very high values of
$R^2$.

\begin{tabular}{l|l|l|l|l}
Data     & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
Exponent & $2.629$ & $2.493$ & $1.814$ & $2.061$
\end{tabular}

In line he two algorithms have almost identical
compression ratios and their fidelity is about the same both subjectively and
based on levenshtein distance. For an example of the text generated on one of
the gen-n datasets, see the appendix.

The runtimes are noticeably worse than for the rep-n dataset. The similarity
algorithm runs in roughly $n^{2.5}$ time while cluster runs in roughly $n^2$
time.

Changing epsilon in this case had a large effect on both the compression ratio
and the fidelity of the final output string.

\subsubsection{Summary}

Both of the algorithms were able to achieve great additional compression ratios
on top of \Sequitur of up to about 40 percent for the highly structured rep-n
data, and 30 percent for the more loosely structured gen-n data without using
the grammar reducer. The fidelity between the original and lossified strings
wasn't great as measured by levenshtein distance, but the texture in both the
rep-n and gen-n datasets was preserved well. Decreasing the $\epsilon$
threshold improved fidelity greatly for larger very structured data at little
cost to the compression ratio, but hurt the compression ratio a lot for less
structured data.

The differences between the two algorithms lie almost entirely in their
runtime. Both algorithms perform much worse on the less structued gen-n data,
we believe because simplifications are made more incrementally and the grammar
does not get dramatically simpler after a few iterations of our algorithm like
it does for the rep-n data.  Unfortunately the \Similarity algorithm tends to
consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
the \Cluster algorithm. Because of this we dismiss the \Similarity algorithm as
unscalable and do not test it further.

@@

Although the grammars generated by \Sequitur are irreducible, our lossifiers
produce grammars which could potentially be compressed further (losslessly)
using Reduction Rules without changing the output string. To test the cost and
effectiveness of applying reduction rules, we looked at the rep-n and gen-n
test cases again, this time focusing on the clustering algorithm for $\epsilon
= .2$. This time we give the ratio on top of the simplified grammar produced by
the clustering algorithm, and the time taken to run just the reduction rules.

When plotted on a log-log scale, the runtime for our straightforward
implementation of the reduction rules appears to be superpolynomial. While
modest additional gains of about 30 percent are possible for highly
self-similar data, very few additional gains are possible for more diverse
datasets such as gen-n. This appears to be because the only simplifications
that are possible on diverse datasets don't lead to the wide-ranging repetitive
substitutions which allow reduction rules to be allowed.  The simplifications
in gen-n tend to be more local.

Because we were unable to get reduction rules to scale to larger datasets, we
will not consider them further.

@@

Though we were ultimately not able to come up with a good way to quantify how
much of the structure of a text was preserved, we have included the final
output strings we obtain after running the \Cluster algorithm with $\epsilon
= .4$ in the Appendices. We believe these tests serve as a good (subjective)
summary of the strengths and weaknesses of our algorithm in compressing data
while preserving its character.

The first piece of text is from the gen-n dataset again, the first 256000
characters of the book of Genesis. As listed in table \ref{simclus_g_t}, the
compressed grammar was size 5540 compared with \Sequitur's size of 8060.
Without knowledge of English spelling, the \Cluster algorithm produces output
that is hard to read. However, key words are clearly recognizable, especially
important words such as `God', and the overall structure in the book of Genesis
is retained.

The second piece of text is one of the entries to this year's International
Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
.  The size of the lossified grammar is 1576 compared with the original size of
1973. The intricate whitespacing was lost, but the clear delineation into
sections was retained, key words such as `define' and `for' are very visible,
and the texture of the obfuscated C is very visible.

>>>>

\begin{enumerate}
  \item Infer a CFG representation of a stream (we use \Sequitur).
  \item Simplify the CFG, introducing loss.
  \item Encode and output the CFG.
\end{enumerate}

@@

In this section, we discuss some of the details of our implementation for
lossifying and reducing CFGs.

@@

\subsubsection{Methodology}

@@

For our first suite of tests, we compared the effectiveness of the
\Similarity and \Cluster algorithms. We might expect \Similarity to
achieve better quality at the same compression ratio since it updates the state
of the grammar between each variable replacement, on the other hand since it
must update the grammar it has the potential to run much more slowly than
\Cluster.

In table \ref{simclus_r_t} we report the results on running \Sequitur and
\Cluster, without the grammar reductions, on the rep-n datasets. In table
\ref{simclus_g_t} we do the same for the gen-n datasets.
In these tests
we were looking to evaluate the relative fidelity, compression ratios, and runtime 
of the two algorithms.

In the tables, Time is the
wall clock running as measured on a 2.26 Ghz Intel Core 2 Duo Macbook Pro.
Ratio is the relative size of the grammar produced compared with \Sequitur,
where the size of the grammar is the sum of the lengths of the rhs's of the
rules. Levdistance is the levenshtein distance between the output of the
lossified grammar and the original input string, it is a kind of measure of the
fidelity of our algorithms but doesn't take into account how well patterns were
preserved. For the rep-n tests, proportion is the proportion of the final
string produced that was `*'. The closer to 1/30~0.033, the more accurate the
lossifier was in recreating the original texture of the input.

\subsubsection{Result Summary}

In both datasets, the 
compression ratios achieved for both algorithms are roughly equal given the
same value for $\epsilon$, so it is fair to compare the algorithms for the same
values of $\epsilon$. 
From plotting the data, we see that runtime of the algorithms are roughly
polynomial in the size of the data, so we can perform a linear regression on
the log-log transform of the data to find the exponents $\alpha$ for the
runtime $r(n)=n^\alpha$ where $n$ is the length of the input string. In the
charts below, all of the linear regressions had $R^2$ values of over $0.98$.

\begin{tabular}{l|l|l|l|l}
rep-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
Exponent      & $1.859$ & $2.026$ & $1.306$ & $1.305$
\end{tabular}

\begin{tabular}{l|l|l|l|l}
gen-n runtime & Sim: $\epsilon=.4$ & Sim: $\epsilon=.2$ & Clust: $\epsilon = .4$ & Clust: $\epsilon = .2$ \\
Exponent      & $2.629$ & $2.493$ & $1.814$ & $2.061$
\end{tabular}

In general, \Cluster has a runtime which is slightly better than quadratic,
while \Similarity has a runtime which is slightly worse than quadratic.
Differant values of $\epsilon$ have comparatively smaller effects on the
runtime, and both algorithms performed worse on the gen-n dataset. We think
this is because there are more tiny substitutions one can perform on
the gen-n dataset, whereas the rep-n dataset allows for much larger scale
substitutions which eliminate the need for many small substitutions.
As expected, \Cluster has a significantly better runtime than \Similarity,
but it is interesting that both algorithms performed much better than our
worst case analysis, we believe it is for the same reasons we proposed
in our runtime analysis section.

Across all of the tests, the leveshtein
distances were comparable for both algorithms. Though they differed in
individual cases, no algorithm did consistently better than the other.
In general the levenshtein distances were high, except in the $\epsilon=.2$
case for the gen-n data where very little compression was possible at all.
We believe most of the distance could be explained by
the fact that the algorithms could drastically shorten
the strings, since it always replaces variables with \emph{smaller} variables
as discussed earlier.
The proportion of '*'s in the rep-n input
was usually the right order of magnitude, with both algorithms making big
mistakes occassionally. In most cases, we felt that the two algorithms maintained similarly
acceptable fidelity levels in recreating the texture of the original rep-n
input file, and we discuss this subjective impression further in the
Larger Tests section.

The actual compression ratio was respectable in most tests, with both \Similarity and \Cluster
coming in at nearly half the size of the \Sequitur CFG for the rep-n data,
and 70 percent for the gen-n data. A notable exception was
the gen-n test with $\epsilon=.2$, where $\epsilon$ was so small that barely
any substitutions were possible. 
However, setting higher values
for $\epsilon$ had diminishing returns for compression for larger
file sizes, at the cost of drastically lowering the fidelity of the output.
Data with differing amuounts of repetition and noise
appear to require different values of $\epsilon$ for the optimal fidelity/compression ratio.

In summary, both of the algorithms were able to achieve moderate additional compression ratios
on top of \Sequitur. The fidelity between the original and lossified strings
wasn't great as measured by levenshtein distance, but we think the texture in both the
rep-n and gen-n datasets was preserved well. Both fidelity and compression ratio were
very sensitive to $\epsilon$ depending on the kind of data.
The differences between the two algorithms lie almost entirely in their
runtime. Empirically the \Similarity algorithm tends to
consistently incur an overhead of about $O(\sqrt{n})$ on top of the runtime for
the \Cluster algorithm. Because of this we dismiss the \Similarity algorithm as
unscalable and do not test it further.

@@

Although the grammars generated by \Sequitur are irreducible, our lossifiers
produce grammars which could potentially be compressed further (losslessly)
using Reduction Rules without changing the output string. To test these rules, 
we looked at the rep-n and gen-n
test cases again, this time focusing on the clustering algorithm for $\epsilon
= .2$. We report our results in table \ref{red_t}, where
we give the ratio on top of the simplified grammar produced by
the clustering algorithm, and the time taken to run just the reduction rules.

When plotted on a log-log scale, the runtime for our straightforward
implementation of the reduction rules appears to be superpolynomial. While
modest additional gains of about 30 percent are possible for highly
self-similar data, very few additional gains are possible for more diverse
datasets such as gen-n. This appears to be because the only simplifications
that are possible on diverse datasets don't lead to the wide-ranging repetitive
substitutions which allow reduction rules to be allowed.

Because we were unable to get our implementation of
reduction rules to scale to larger datasets, we
will not consider them further.

@@

Though we were ultimately not able to come up with a good way to quantify how
much of the structure of a text was preserved, we have included some final
output strings obtained after running the \Cluster algorithm with $\epsilon
= .4$ in the Appendices. We believe these tests serve as a good (subjective)
summary of the strengths and weaknesses of our algorithm in compressing data
while preserving its character.

The first piece of text is from the gen-n dataset again, the first 256000
characters of the book of Genesis. As listed in table \ref{simclus_g_t}, the
compressed grammar was size 5540 compared with \Sequitur's size of 8060.
Without knowledge of English spelling, the \Cluster algorithm produces output
that is hard to read. However, key words are clearly recognizable, especially
important words such as `God', and the overall division of the book into
verses was retained.

The second piece of text is one of the entries to this year's International
Obfuscated C code contest, at \url{http://www.ioccc.org/2012/deckmyn/deckmyn.c}
.  The size of the lossified grammar is 1576 compared with the original size of
1973. The intricate whitespacing was lost, but the clear delineation into
sections was kept, key words such as `define' and `for' are very visible,
and the texture of the obfuscated C is very obvious.
